A Coq Formalization of Stable Sorts, Radix Sort, and the Burrows-Wheeler Transform

### 2019-10-20 draft_paper/draft_paper.tex
Fix a few spelling errors

This intuition is related to the following lemma.
\begin{restatable*}[lexsort\_rots\_hdsort]{lemma}{lexsrotrotshdsort}
  For all lists $l$,
  \begin{equation*}
    (\var{hdsort} \comp
    \var{map}~\var{rrot})~(\var{lexsort}~(\var{rots}~l)) =
    \var{lexsort}~(\var{rots}~l)
  \end{equation*}
\end{restatable*}
This lemma says that if we take the last column of the sorted rotation
matrix and move it to the front, and then sort on it, we produce the
same sorted rotation matrix that we started with. This lemma means
that when we repeatedly prepend $s$ and sort by it, we are preserving
the rest of the sorted rotation matrix, even the parts we have yet to
reconstruct.

@@

The specificiation for \var{recreate} is that it recreates the first
$j$ columns of the sorted rotation matrix. For this we need to define
\var{cols}.
\begin{lstlisting}
Definition cols (j : nat)
  : list (list A) -> list (list A)
  := map (firstn j).
\end{lstlisting}

@@

\lexsrotrotshdsort
\begin{proof}
  By \cref{thm:sort_ord_unique}, any two sorts on an order are
  equivalent, and we can replace \var{lexsort} with \var{radixsort}.
  \begin{align*}
       &(\var{hdsort} \comp \var{map}~\var{rrot})~(\var{lexsort}~(\var{rots}~l)) \\
    ={}&(\var{hdsort} \comp \var{map}~\var{rrot})~((\var{hdsort} \comp \var{map}~\var{rrot})^n~(\var{rots}~l)) \\
    ={}&(\var{hdsort} \comp \var{map}~\var{rrot})^{n+1}~(\var{rots}~l) \\
    ={}&(\var{hdsort} \comp \var{map}~\var{rrot})^n~((\var{hdsort} \comp \var{map}~\var{rrot})~(\var{rots}~l)) \\
    ={}&\var{lexsort}~((\var{hdsort} \comp \var{map}~\var{rrot})~(\var{rots}~l)) \\
    ={}&\var{lexsort}~((\var{hdsort} \comp \var{rrot})~(\var{rots}~l)) \\
    ={}&\var{lexsort}~(\var{rots}~l)
  \end{align*}
  The last step follows because the function $\var{hdsort} \comp
  \var{rrot}$ permutes the rotation matrix right before it is sorted
  by \var{lexsrot}, so it doesn't change the final result.
\end{proof}

@@

We can define the image and preimage of an index under a permutation
function.
\begin{lstlisting}
Definition image (p : list nat) i := findIndex i p.
Definition preimage (p : list nat) i := nth i p 0.
\end{lstlisting}
We will use the notation $p(i)$ for $\var{image}~i$ and $p^{-1}(i)$
for $\var{preimage}~i$. It should be clear from context whether $p(x)$
denotes $\var{image}~x$ or $\var{apply}~x$, and when it is not we will
use \var{image} and \var{apply} explicitlity.

>>>>

This intuition is related to the following lemma.
\begin{restatable*}[lexsort\_rots\_hdsort]{lemma}{lexsortrotshdsort}
  For all lists $l$,
  \begin{equation*}
    (\var{hdsort} \comp
    \var{map}~\var{rrot})~(\var{lexsort}~(\var{rots}~l)) =
    \var{lexsort}~(\var{rots}~l)
  \end{equation*}
\end{restatable*}
This lemma says that if we take the last column of the sorted rotation
matrix and move it to the front, and then sort on it, we produce the
same sorted rotation matrix that we started with. This lemma means
that when we repeatedly prepend $s$ and sort by it, we are preserving
the rest of the sorted rotation matrix, even the parts we have yet to
reconstruct.

@@

The specification for \var{recreate} is that it recreates the first
$j$ columns of the sorted rotation matrix. For this we need to define
\var{cols}.
\begin{lstlisting}
Definition cols (j : nat)
  : list (list A) -> list (list A)
  := map (firstn j).
\end{lstlisting}

@@

\lexsortrotshdsort
\begin{proof}
  By \cref{thm:sort_ord_unique}, any two sorts on an order are
  equivalent, and we can replace \var{lexsort} with \var{radixsort}.
  \begin{align*}
       &(\var{hdsort} \comp \var{map}~\var{rrot})~(\var{lexsort}~(\var{rots}~l)) \\
    ={}&(\var{hdsort} \comp \var{map}~\var{rrot})~((\var{hdsort} \comp \var{map}~\var{rrot})^n~(\var{rots}~l)) \\
    ={}&(\var{hdsort} \comp \var{map}~\var{rrot})^{n+1}~(\var{rots}~l) \\
    ={}&(\var{hdsort} \comp \var{map}~\var{rrot})^n~((\var{hdsort} \comp \var{map}~\var{rrot})~(\var{rots}~l)) \\
    ={}&\var{lexsort}~((\var{hdsort} \comp \var{map}~\var{rrot})~(\var{rots}~l)) \\
    ={}&\var{lexsort}~((\var{hdsort} \comp \var{rrot})~(\var{rots}~l)) \\
    ={}&\var{lexsort}~(\var{rots}~l)
  \end{align*}
  The last step follows because the function $\var{hdsort} \comp
  \var{rrot}$ permutes the rotation matrix right before it is sorted
  by \var{lexsort}, so it doesn't change the final result.
\end{proof}

@@

We can define the image and preimage of an index under a permutation
function.
\begin{lstlisting}
Definition image (p : list nat) i := findIndex i p.
Definition preimage (p : list nat) i := nth i p 0.
\end{lstlisting}
We will use the notation $p(i)$ for $\var{image}~i$ and $p^{-1}(i)$
for $\var{preimage}~i$. It should be clear from context whether $p(x)$
denotes $\var{image}~x$ or $\var{apply}~x$, and when it is not we will
use \var{image} and \var{apply} explicitly.
